---
title: "Is there a beneficial side of COVID?"
author: "Chee Kian Teoh, David Zhao, Harsahib Grewal"
date: "November 12th, 2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

List your group members, including their student numbers, here:

-   David Zhao (169101599)
-   Chee Kian Teoh (210984770)
-   Harsahib Grewal (\#########)

```{r setup, include=FALSE}
# echo = FALSE will set the Rmd to *not* show the R code. Don't change this.
# You may change the default figure width and figure height as you please.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 6)

# Put any libraries that you need to load here.
# DO NOT PUT "install.packages()" IN AN RMD FILE!!!
library(tidyverse)
library(arrow)
library(patchwork)
library(ggExtra)
```

# Abstract

We are interested in exploring how COVID-19 have impacted climate change by analyzing three datasets: `covid_2020_original.parquet` (COVID), `ice_extent_daily.parquet` (Ice Extent), and `cyclones_data.parquet` (Hurricane). To examine relationships among these datasets, we conduct statistical analyses (summary statistic) and implement data visualization (plotting graphs) to identify trends and possible correlations. Surprisingly, our findings indicate that COVID-19 did not alleviate climate conditions.

# Introduction

Climate change is a topic of extensive study and recently more challenges have arisen for environmental and public health, economic stability[], and global ecosystems[\^footnote]. As the climate continues to change because of natural events and human activities, the consequences are becoming more and more visible. The analysis looks at some of these impacts by getting data on natural events and public health and connecting them to see any emerging patterns that can lead to the answer of if they have any relation to the environmental change over recent decades.

In this report, some key indicators of climate change versus global response factors are explored through exploratory analysis. The dataset we use look at the strength, and frequency of hurricanes in the Atlantic and North Pacific basins; polar ice extent in the Arctic and Antarctic regions; and COVID-19 reported cases around the world during the year 2020. Some techniques used in this analysis are data visualization and statistical explorations of any trends and possible correlations (whether high or low) within the data.

The goal of this analysis is to learn patterns in these datasets and see the finer relation among these factors that may provide insight into what is affecting climate change. There are several limitations to the data here mainly around time and geographical locations, but this report still tries to show the relationships of climate events and their outcomes.

By the end of this report, we will explain these correlations and provide a conclusion supported with evidence of the current impact of climate change.

# Data Description

## \<\<Hurricane\>\>

```{r load_data1}

# hurricane
cyclones_data_original <- read_parquet("cyclones_data.parquet")

cyclones_data <- cyclones_data_original |> 
  group_by(Basin, ObservYear) |>
  mutate(AvgMaxWindPerYear = mean(max_wind, na.rm = TRUE)) |>
  select(Basin, ObservYear, AvgMaxWindPerYear) |>
  distinct() 

cyclones_data
```

The cyclones dataset shows data on hurricane activity across the Atlantic and North Pacific basins. Some important columns include cyclone strength (maximum wind speed, and category), cyclone duration, and basin of origin. More columns had similar attributes which were grouped such as geographical coordinates (latitude and longitude) of the cyclones at recorded times and dates of occurrence. This dataset allows us to analyze hurricane frequency, strength, and duration trends across different regions.

In order to clean the data, we:

\- Remove any records with missing values in key columns such as wind speed and pressure to ensure accuracy in cyclone strength measurements.

\- Standardize date and time formats and convert them to datetime types where applicable.

\- Remove any duplicate records based on unique cyclone identifiers.

## \<\<Covid\>\>

```{r load_data2}
# Put in your code to load in the data set, along with any
# necessary cleaning beyond what was done in Part 1

# Reminder: do not print your data to the screen unless it's
# completely necessary

covid_2020_original <- read_parquet("covid_2020_original.parquet")

covid_2020 <- covid_2020_original |>
  select(
    "date", 
    "total_cases", 
    "total_cases_per_million", 
    "new_cases",
    "new_cases_per_million", 
    "stringency_index",
  )
covid_2020
```

The COVID dataset reports total COVID-19 cases from 2020 to 2024 looking at most of the countries. Some important columns are the country name, total reported cases (new and old), deaths, population, stringency index and other demographic details relevant to the country's deaths to cases ratio. This dataset has a lot of empty data which would've proven useful to find some relations (ICU/hospital admissions and patients, COVID tests taken, and COVID vaccinations taken). Regardless of the missing material, this dataset allows us to explore the pandemic's global impact in 2020.

In order to clean the data, we:

\- Remove all empty columns with missing data values

\- Remove records with missing or zero values in the total cases and population columns.

\- Standardize country names to ensure a common trend across the data set.

\- Convert any non-numeric data types in the reported cases and population columns to numeric for easier calculations.

## \<\<Ice extent\>\>

```{r load_data3}
ice_extent_daily_original <- read_parquet("ice_extent_daily.parquet")

ice_extent_daily <- ice_extent_daily_original |>
  pivot_wider(
    names_from = region,
    values_from = ice_extent
  )
ice_extent_daily
```

This dataset contains yearly data on ice extent in both the Arctic and Antarctic regions. Some important columns were year, region (Arctic or Antarctic), and ice extent measurements (km\^2). This dataset shows polar ice coverage which is a significant cause of climate change, and this allows us to see the trend analysis over time to see potential reductions in ice extent.

In order to clean the data, we:

\- Clean any missing values in the ice extent column to avoid gaps in yearly trend analysis.

\- Convert the year column to datetime for consistency.

\- Check for any duplicate records by year and region to maintain unique records.

## Combining the Data

For the combination of the ice_extent dataset and covid_2020 dataset we were interested in combining both tables by their "date" column so we can analyze and see the correlation between the melting ice and COVID-19 stringency index (a measurement of how strict government policies were throughout the COVID-19 pandemic).

To achieve the combination we used "left_join"[^4] to combine the ice_extent dataset into the COVID_2020 dataset. The reasoning behind choosing to combine with a left_join is because the ice_extent dataset has a wider range of data collected going from 1980 to 2023 whereas covid_2020 only had data from 2020 to 2024. Because of this we wanted to combine according to the same year while showing data from previous year in the ice_extent dataset to reflect if there is a correlation between the two.

# Exploratory Data Analysis

We explored many aspects of the data, but will demonstrate three. COVID did not contribute a significant change in \<\<the wind speed of hurricanes\>\>, \<\<The frequency and the type of hurricane occurring per year\>\>, and \<<ice melting speed per year>\>

The first aspect that we found interesting is shown in \@ref(fig:insight1). The insight should be specific to the data shown, not a general statement beyond the data (leave that for the conclusion).

```{r insight1, fig.cap="The trend of hurricane average max wind per year for both Atlantic and Eastern Pacific from 1990 to 2022"}
# This is an example of how you can control figures and captions in
# an R chunk. Note that you can reference figures using:
# \@ref(fig:insight1), where "insight1" is the label of this code
# chunk (the first bit of text after the "r" in "```{r label, options...}")
cyclones_data |>
  filter(Basin != "CP")|>
  ggplot() +
    aes(x = ObservYear, y = AvgMaxWindPerYear, colour = Basin) +
    geom_point() +
    geom_smooth(se = FALSE, method = 'loess', formula = 'y ~ x') +
    xlim(1990, 2022) +
    labs(x = "Observe Year", y = "Average Max Wind (knots)", title = "Hurricane Average Max Wind for Both Atlantic And Eastern Pacific", subtitle = "From 1990 To 2022")
```

The data come from \<\<cyclones_data\>\> and describe the trend for Atlantic and Eastern Pacific from 1990 to 2022. They seem having complete opposite slope and it is not quite obvious to see a change between 2019 to 2023 (starting and end time for COVID) for individual locations. Additionally, A big factor on speed and the frequency of spearing is heavily dependent on the current temperature. If COVID indeed reduces the temperature of Earth due to less human activity, there should be a difference in trend compare to the last 10 years. This shows that there is no significant effect in terms of hurricane speed.

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

```{r summary_stats}
# we interested in the relationship between covid19 and the effect of ice extent ice extent
# covid19 and ice extent
# covid19 human activities related variables: new_cases, new_deaths,
# ice_extent: date, region, ice_extent

ice_extent_daily_original <- read_parquet("ice_extent_daily.parquet")

ice_extent_daily <- ice_extent_daily_original |>
  pivot_wider(
    names_from = region,
    values_from = ice_extent
  )

# Joined both covid_2020 and ice_extent_daily together by date
covid_2020_ice_extent <- ice_extent_daily |> 
  left_join(
    covid_2020,
    by = join_by(date)
  ) 

# We interest in the sudden spike of stringency index overall
# measure of restriction and lockdown: https://ourworldindata.org/covid-stringency-index
# The stringency index typically measures the strictness of government policies and interventions during the COVID-19 pandemic, such as lockdowns, travel bans, and restrictions on businesses and social gatherings. A higher value indicates stricter measures, while a lower value indicates more relaxed or fewer restrictions.
monthly_ice_extent_and_stringency <- covid_2020_ice_extent |>
  pivot_longer(
    cols = c(Antarctic, Arctic),
    names_to = "region",
    values_to = "ice_extent"
  ) |>
  group_by(year, month) |>
  mutate(
    stringency_index_monthly_avg = mean(stringency_index, na.rm = TRUE),
    average_ice_melt = mean(ice_extent, na.rm = TRUE)
  ) |>
  ungroup()

# correlation
monthly_ice_extent_and_stringency |>
  summarise(
    correlation = round(cor(
      stringency_index_monthly_avg, 
      average_ice_melt, 
      use = "complete.obs"
    ) , 3)
  ) |>
  rename("correlation between average ice melting monthly and the strictness of COVID" = "correlation")

# correlation graph
cor_value <- cor(monthly_ice_extent_and_stringency$stringency_index_monthly_avg, 
                 monthly_ice_extent_and_stringency$average_ice_melt, 
                 use = "complete.obs") 

p <- ggplot(monthly_ice_extent_and_stringency, aes(x = stringency_index_monthly_avg, y = average_ice_melt)) +
  geom_point(alpha = 0.6, na.rm = TRUE) +  
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) + 
  labs(
    title = "Correlation between Monthly Average Ice Extent\n and COVID-19 Stringency Index over time",
    subtitle = str_c("Correlation Coefficient: ", round(cor_value, 3)),
    x = "COVID-19 Stringency Index",
    y = "Average Ice Extent (unit)"
  ) +
  theme_minimal()

p

# Finding: An increase in the average stringency index associates with very little or no increase of average ice extent.

```

The summary statistic is the correlation coefficient between the COVID-19 stringency index (a measurement of how strict government policies were throughout the COVID-19 pandemic) and the average ice melting area(km\^2) every month. This correlation is very weak. We can conclude that there is not correlation between the two variables. However, this might be caused by other factors. Maybe the effect from lock down would not occur immediately. To find out, lets make some plots according to time and analyse the trend.

The next insight that we found is shown in \@ref(fig:insight2).

```{r insight2, fig.height=4, fig.width=6, fig.cap="The frequency of different hurricane categories in each year from 2014 to 2022. The hurricane categories represent by Saffir-Simpson scale <- reference? "}
# This figure will have a height of 4 and a width of 6.
# Feel free to change this, and to apply different sizes
# to the other figures you create.

# Put in your code to load in the data set, along with any
# necessary cleaning beyond what was done in Part 1

# Reminder: do not print your data to the screen unless it's completely necessary


# Hurricane facet by year
# x = categories
# y = frequencies (how many times each categories of hurricane happens in that year)
cyclone_category_frequency_from_2015_to_2022 <- cyclones_data_original |>
  filter(ObservYear >= 2014 & ObservYear <= 2022) |>
  ggplot(aes(x = category)) +
  geom_bar() + 
  facet_wrap(~ObservYear)

cyclone_category_frequency_from_2015_to_2022
```

The data come from \<\<cyclones_data_original\>\> and it is use the analyse the category of hurricanes. Across from different graphs for years, all of them keeps a right skewed trend. Looking into the years(2019 to 2022) when COVID is present, The number of stronger Hurricanes are appearing less then the years before. This could be an effect of COVID. However, this also could be other factors in the world because in the previous year there are also similar changes.

Finally, \@ref(fig:insight3) shows ...

```{r insight3, fig.height=5, fig.width=7, fig.cap="Normalized Ice Extent and COVID-19 Stringency Index from 2016 to 2023"}

# Note: Please pick between one of the graph below:
# 1. Normalized Ice Extent and COVID-19 Stringency Index (2016-2023)
# 2. Correlation between Monthly Average Ice Extent\n and COVID-19 Stringency Index over time

# 1. Comparison graph between ice extent and stringency index
# Pros: Different plot type, can see more comparison together
# Cons: Complex than the correlation graph(2), harder to understand the effect
range_y_stringency_index <- max(monthly_ice_extent_and_stringency$stringency_index, na.rm = TRUE) - min(monthly_ice_extent_and_stringency$stringency_index, na.rm = TRUE)
range_y_avg_ice_melt <- max(monthly_ice_extent_and_stringency$average_ice_melt, na.rm = TRUE) - min(monthly_ice_extent_and_stringency$average_ice_melt, na.rm = TRUE)

scaling_factor <- range_y_stringency_index / range_y_avg_ice_melt

year_lines <- seq.Date(
  from = as.Date("2016-01-01"), # change start date based on your data
  to = as.Date("2024-01-01"),   # change end date based on your data
  by = "year"
)


monthly_ice_extent_and_stringency |>
  mutate(
    # Normalize ice extent data to a 0-1 scale
    normalized_ice_extent = (
      average_ice_melt - min(average_ice_melt, na.rm = TRUE)
    ) / (
      max(average_ice_melt, na.rm = TRUE) - min(average_ice_melt, na.rm = TRUE)
    ),
    
    # Normalize COVID-19 stringency index to a 0-1 scale
    normalized_stringency_index = stringency_index_monthly_avg / 100
  ) |>
  ggplot() +
  aes(x = date) +
  geom_area(aes(y = normalized_stringency_index), fill="#69b3a2", alpha=0.4, na.rm = TRUE) +
  geom_line(aes(y = normalized_stringency_index, color = "COVID-19 Stringency Index"), linewidth=0.5, na.rm = TRUE) +
  geom_line(aes(y = normalized_ice_extent, color="Ice Extent"), linewidth=0.5, na.rm = TRUE) +
  geom_vline(xintercept = as.numeric(year_lines), color = "black", linetype = "dashed", alpha=0.4) +
  scale_x_date(
    limits = as.Date(c("2016-01-01","2023-01-01")),
    date_breaks = "1 years",
    date_labels = "%Y (%b)",
  ) +
  labs(
    title="Normalized Ice Extent and COVID-19 Stringency Index (2016-2023)",
    subtitle = "Monthly trends, both scaled to 0-1 to observe potential correlations",
    x = NULL,
    y = "Standardized Ice Extent and Stringency Index (0-1)"
    
  ) +
  scale_color_manual(
    name = "Legend",                       
    values = c("Ice Extent" = "blue",     
               "COVID-19 Stringency Index" = "red") 
  ) +
  theme_grey() +
  theme(legend.position = "top")


```

This graph is generated by combining information form ice extent and the strictness for COVID regulations defined by the law at the time. The bar graph above shows the area of ice melted every year in km\^2. The green bar plot is showing the strictness of law due to the COVID pandemic. Strict polices can create less human activity which allow us observe the difference for ice melting area when there is less human activity. However, this is not the case. When the enforcement of new laws regarding the pandemic increases, the ice melting area didn't significantly change as the shapes of distribution on the years when COVID hit is relatively the same as the past years.

# Conclusion and Future Work

Overall, we found \<<COVID-19 did not have significant drawback on global warming>\>[\^].

We found that COVID did not change the climate condition earth because there is not significant changes in speed of Hurricanes, frequency of Hurricane occurring, and ice bergs melting speed. All these factors that are heavy dependent on the temperature condition are not getting influenced by less human activity during the pandemic. Additional, the correctional generated from ice melting speed and the strictness of law for lock down also reflect there is not relationship between climate conditions nd COVID19.

The next steps in this analysis are collecting more useful data sets such as carbon emission level and earth temperature. By analyzing the relationship between COVID and these data set, we can get more significant insights about this topic.

**The limitations of this analysis are as follows.**

-   There are lots of other climate conditions that occurs during COVID time period which is a significant challenge towards success. To minimize this issue, Collect more data sets to make more accurate analysis and understand the relationship between each other and come up with a more accurate and effective conclusion.

# References

I am not strict about MLA or APA style or anything like that. For this report, I would much rather have your citations be easy to match to your insights.

The easiest way is to use Rmd's [footnote](https://bookdown.org/yihui/rmarkdown/markdown-syntax.html#inline-formatting) syntax. This will put a number beside the word where the footnote appears, and the full text of the footnote at the bottom of the page (pdf) or end of the document (html). The syntax is:[^1], where I suggest that you put in something like this[^2] to make references for this assignment.

[^1]: The relevance to the insight is ... . \
    From Climate Change Indicators and Impacts Worsened 2020: World Meteorological Organization, published on 19 April 2021, URL: <https://wmo.int/news/media-centre/climate-change-indicators-and-impacts-worsened-2020>

[^2]: The relevance to the insight is ... . From \<<name of source and name of article>\>, published on \<<date>\>, url: \<<link to page>\>

[^4]: Learn about the concept of left join, right join and inner join. From Joining data frames with dplyr, url: https://lindsaydbrin.github.io/CREATE_R_Workshop/Lesson_-_dplyr_join.html

make sure that you provide a sentence to describe the source and it's relevance when you cite it - don't just add citations to the end of a sentence (this is common practice in academia, but I want to know that your citations are directly relevant for this assignmnet).
